Multi-fluids for Representing Sub-filter-scale Convection

### 2020-05-04 submit2/manuscript.tex
Nearly ready to resubmit. A better pressure perturbation model is used.

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%
\addeditor{HW}
\addeditor{WM}

@@

\citeA{TWV+18} proposed multi-fluid modelling of convection in order
to treat non-equilibrium and net mass transport \remove[HW]{correctly and} consistently.
\add[HW]{The multi-fluids approach has an advantage over ADHOC and CLUBB in solving prognostic equations for first-order moments directly (mean properties of eg. updraft and environment) rather than solving prognostic equations for higher-order moments of the whole fluid in order to diagnose means of updrafts and the environment. Solving directly for sub-fluid means should be less sensitive to closure assumptions.}
\remove{with the same numerics and physics to prognose properties of separate parts of the fluid, for example updrafts, the stable environment and downdrafts dynamics}. \citeA{Yano14} derived similar
equations \remove[HW]{for multiple segments} and showed how this is a generalisation
of the mass-flux formulation. The \citeA{TKP+18} ``extended eddy
diffusivity mass flux'' or extended EDMF scheme \remove[HW]{also} uses similar
equations but \citeA{TKP+18}
\change[HW]{didn't propose a suitable numerical method for solving the equations.}
{only solve the equations in a single column without coupling to the continuity equation of a dynamical core.}
\change[HW]{Multi-fluids promises big improvements
in the representation of sub-grid-scale convection and unification
with the representation of turbulence, particularly at the grey zone.
However, there is a long way to go, particularly as multi-fluids}
{The disadvantage of the multi-fluids approach is that it} requires
a dedicated model rather than being a stand alone parameterisation
for a dynamical core. \citeA{TEB19} and \citeA{WM19} presented solutions
of \change[HW]{the multi}{two}-fluid equations
\change[HW]{assuming two fluids that share}{using} the same pressure
\add[HW]{for both fluids}.
\remove[HW]{(or that pressure differences are predominantly due to drag
between the fluids)}. \citeA{TEB19} showed that these equations \remove[HW]{(without
drag)} are in fact unstable and can be stabilised by diffusion of vertical
velocity whereas \citeA{WM19} stabilised the equations with diffusion
and drag between fluids. Neither of these stabilisation techniques
are acceptable;
\add[HW]{sufficient diffusion to stabilise the multi-fluid equations may smooth out real features of the flow and the diffusion between fluids and drag used by \protect\citeA{WM19} meant that both fluids tended to move together rather than through each other.}
In this paper we show that there are significant
and sustained pressure differences between the fluids (section \ref{sec:results}) \add[HW]{and that the use of separate pressures stabilises the equations}.

@@

Many of the challenges of representing convection with the multi-fluid
Navier-Stokes equations \cite<e.g. those described in>{WM19} carry
over to the Boussinesq equations \add[HW]{because} convection is buoyancy dominated
and close to incompressible. The resolved test cases presented in
section \ref{sec:results} give very similar results for Navier-Stokes
and Buossinesq equations. Therefore\change[WM]{ for simplicity we focus on the}{, the}
\remove[HW]{multi-fluid} Boussinesq equations without background stratification \add[WM]{are chosen for simplicity}:
\note[HW]{This is additional material until the start of 2.1}
\begin{eqnarray}
\frac{\partial \mathbf{u}}{\partial t} + \nabla\cdot(\mathbf{u}\mathbf{u})
+\nabla P & = & b \mathbf{k}+\nu\nabla^{2}\mathbf{u}
\label{eq:singleMom}
\\
\frac{\partial b}{\partial t} + \nabla\cdot(\mathbf{u}b)
 & = & \alpha\nabla^{2}b
\label{eq:singleb}
\\
\nabla\cdot\mathbf{u} & = & 0
\label{eq:singleDivFree}
\end{eqnarray}
where variables are defined in table \ref{tab:defns}.

@@

The single-fluid Boussinesq equations will be conditionally averaged following \citeA{TWV+18} to give the multi-fluid Boussinesq equations. Lagrangian labels for each fluid $i$ are defined as $I_i=1$ in fluid $i$ and zero elsewhere. 
For the derivation of the conditionally averaged equations \change[WM]{we will not consider relabelling of fluid}{the relabelling of fluids will not be considered,} \add[HW]{and so} $I_i$ satisfies:
\begin{equation}
\frac{DI_i}{Dt}=0.
\label{eq:LagrangianLabel}
\end{equation}
\change[WM]{Applying conditional volume average to $I_i$ we define the volume fraction of fluid $i$ and the velocity and buoyancy of fluid $i$}{By applying a conditional volume average (denoted by the overbar), the fluid $i$ volume fractions, velocities and buoyancies are defined by}:
\begin{eqnarray}
\sigma_i &=& \overline{I_i} \\
\sigma_i \mathbf{u}_i &=& \overline{I_i \mathbf{u}} \\
\sigma_i b_i &=& \overline{I_i b}
\label{eq:defineFluidFields}
\end{eqnarray}
which implies
\begin{equation}
\sum_{i}\sigma_{i}  =  1.
\label{eq:sumOne}
\end{equation}
\change[WM]{As we are considering the Boussinesq equations we assume}{As the system is governed by the Boussinesq equations, it is assumed} that variations in density only affect the \change[WM]{buoyancy terms}{definition of the buoyancy terms (table \protect\ref{tab:defns})} so $\mathbf{u}_i$ and $b_i$ are defined without density weighting.

@@

Next \change[WM]{we will consider conditionally averaging the terms of }{the conditional averaging of terms in} the momentum (\ref{eq:singleMom}) and buoyancy (\ref{eq:singleb}) equations\add[WM]{ is considered}. Conditionally averaging the advection terms leads to sub-filter-scale fluxes, $F_{SF}$, since the velocity and buoyancy are not uniform within each fluid at the filter scale:
\begin{eqnarray}
\overline{I_i \nabla \cdot (\mathbf{u} \mathbf{u})} &=& \nabla \cdot (\sigma_i \mathbf{u}_i  \mathbf{u}_i) + F_\text{SF}^{\mathbf{u}_i} \\
\overline{I_i \nabla (\mathbf{u} b)} &=&  \nabla \cdot (\sigma_i \mathbf{u}_i b_i) + F_\text{SF}^{b_i}
\label{eq:filterAdvection}
\end{eqnarray}
The sub-filter-scale fluxes will be absorbed into the diffusion terms for momentum and buoyancy.
%
\note[HW]{Dan, can we do better than this?}
\change[WM]{Lots of}{Many} terms appear when attempting to conditionally average the diffusion terms which we do not know how to parameterise. \change[WM]{Therefore we will use simply}{The conditionally averaged diffusion terms are therefore assumed to behave like}:
\begin{eqnarray}
\overline{I_i \nu \nabla^2 \mathbf{u}} &=& \sigma_i \nu \nabla^2 \mathbf{u}_i \\
\overline{I_i \alpha \nabla^2 b} &=& \sigma_i \alpha \nabla^2b_i .
\end{eqnarray} 
%
\remove[WM]{Unlike \protect\citeA{TWV+18}} \change[WM]{we define separate pressures for each fluid, $P_i$. The pressures are defined to satisfy:}{The pressures for each fluid, $P_i$, are defined to satisfy:}
\begin{equation}
\overline{I_i \nabla P} = \sigma_i \nabla P_i
\end{equation}
\add[WM]{which differs from \protect\citeA{TWV+18} who use a shared fluid pressure.}
%
Combining \change[WM]{these}{the above terms} gives the conditionally averaged momentum and buoyancy equations:
\begin{eqnarray}
\frac{\partial\sigma_i\mathbf{u}_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} \mathbf{u}_{i})
+ \sigma_i \nabla P_{i}
& = &
\sigma_i b_{i}\mathbf{k}
+
\nu\sigma_i\nabla^{2}\mathbf{u}_{i}
+
\sum_{j\ne i}\left\{ \sigma_{j}S_{ji}\mathbf{u}_{ji}^{T}
 - \sigma_{i}\mathbf{D}_{ij}\right\}
\label{eq:momFlux}\\
\frac{\partial\sigma_i b_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} b_{i})
& = &
\alpha\sigma_i\nabla^{2}b_{i}
+
\sum_{j\ne i} \sigma_{j}S_{ji} b_{ji}^{T}.
\label{eq:bFlux}
\end{eqnarray}
These equations include mass transfers between fluids with $\mathbf{u}_{ij}^T$ being the velocity of the fluid transferred from $i$ to $j$ and $b_{ij}^T$ being the buoyancy of the fluid transferred from $i$ to $j$. A drag term, $\mathbf{D}_{ij}$ is also included to parameterise some of the missing terms because it is commonly used when conditional averaging is applied in fields such as multi-phase modelling \cite<eg>{GBB+07}. Sensitivity to $\mathbf{D}_{ij}$ will be investigated in section \ref{sec:results}.

@@

\change[WM]{We will assume that}{It is assumed that the} pressure in each fluid is given by a \remove[HW]{parameterised}
perturbation, $p_{i}$, from the total pressure, $P$:
\begin{equation}
P_{i}=P+p_{i}.
\end{equation}
\add[HW]{In order to derive a suitable form of parameterisation for $p_i$ we will first consider the} \citeA{BN86} \add{model of multi-phase flow.}

@@

\citeA{BN86} proposed a parameterisation for pressure differences
in a multi-phase explosion model that has been widely used in mathematics
and engineering for simulating, \add[HW]{for example}, fluidised beds of granular material \cite<e.g.>{EHM92} and two compressible fluids \cite<e.g.>{SA99}. \add[HW]{This model is not used here but is examined in order to derive a parameterisation of pressure differences between fluids suitable for atmospheric convection.}
\citeA{SA99} wrote transport equations for both the volume fraction,
$\sigma_{i}$, and for the mass fraction,
$\sigma_{i}\rho_{i}$  without transfers between fluids which in our notation are:
\begin{eqnarray}
\frac{\partial\sigma_{i}}{\partial t}+\mathbf{V}\cdot\nabla\sigma_{i} & = & \mu\left(P_{i}-P_{j}\right)
\label{eq:BNcompaction}
\\
\frac{\partial\sigma_{i}\rho_{i}}{\partial t}+\nabla\cdot\left(\sigma_{i}\rho_{i}\mathbf{u}_{i}\right) & = & 0
\label{eq:rhoTransport}
\end{eqnarray}
where $\mathbf{V}$ is the velocity of the interface between fluids,
$\mu$ is the compaction viscosity which controls how quickly the
pressure of each fluid relaxes to the mean pressure and index $j$
refers to the other fluid in a two fluid system. In our Boussinesq
system, $\rho_{i}$ does not appear in the continuity equation. Different
authors make different assumptions about the interface velocity but
if \change[WM]{we assume}{one assumes} that $\mathbf{V}=\sum_{i}\sigma_{i}\mathbf{u}_{i}$ and
$\rho_{i}$ is constant then eqns (\ref{eq:divFree}), (\ref{eq:sumOne}), (\ref{eq:BNcompaction}) and (\ref{eq:rhoTransport})
imply that $\mu p_{i}=-\sigma_{i}^{2}\nabla\cdot\mathbf{u}_{i}$.
\change[WM]{If we assume that $\mathbf{V}=\mathbf{u}_{i}$ for each fluid then
we get $\mu p_{i}=-\sigma_{i}\nabla\cdot\mathbf{u}_{i}$.}{Instead assuming $\mathbf{V}=\mathbf{u}_{i}$ for each fluid results in $\mu p_{i}=-\sigma_{i}\nabla\cdot\mathbf{u}_{i}$.}

@@

\citeA{TEB19} stabilised the multi-fluid equations by adding a diffusion
term $\nu\partial w^{2}/\partial z^{2}$ to the RHS of the $w$ equation.
If a fluid perturbation pressure is set to \change[WM]{$-\nu\partial w/\partial z$}{$p_i=-\nu\partial w/\partial z$}
\add[WM]{then} this is equivalent to the diffusion term in one dimension. However
adding a diffusion term will harm three dimensional simulations of
resolved convection in the free troposphere whereas adding a fluid
perturbation pressure will not. Inspired by both \citeA{BN86} and
\citeA{TEB19}\change[WM]{ we assume that the pressure in each fluid is given
by}{, the pessure in each fluid is chosen to be}:
\begin{equation}
p_{i}=-\gamma\nabla\cdot(\sigma_i \mathbf{u}_{i})
\label{eq:Pi_div}
\end{equation}
where $\gamma$ is a \change[HW]{free coefficient}{compressibility with units $\text{m}^s \text{s}^{-1}$. 
This parameterisation of $p_i$ has $\sigma_i$ inside the divergence in order to ensure that $\sum_i p_i =0$ and therefore that $\sum_i P_i = P$ which makes the equations more striaghtforward to solve.}
\add[HW]{However it should be noted that this parameterisation destroys energy conservation.}

@@

\add[HW]{The transfers between fluids are related to entrainment, $\varepsilon$, and detrainment, $\delta$. Fluid 0 is the environment and fluid 1 is updrafts implying:}
\begin{equation}
S_{01} = w_1\varepsilon, \;\;\;\;\;\;\;
S_{10} = w_0\delta.
\end{equation}
\add[HW]{Results using two formulations of entrainment are presented in section {\protect\ref{sec:results}}. The first is}
\remove[HW]{We consider dry air without turbulence so use}
the dynamic entrainment described by
\citeA{HC51}, \citeA{AK67} and \citeA{DBF+13}:
\begin{equation}
S_{ij}=\begin{cases}
-\nabla\cdot\mathbf{u}_{i} & \text{if }\nabla\cdot\mathbf{u}_{i}<0\\
0 & \text{otherwise.}
\end{cases}
\label{eq:Sdiv}
\end{equation}
\change[HW]{which is necessary to stop air in}{This formulation prevents} one fluid from accumulating when \change[WM]{vertical motion ceases}{a fluid decelerates} (at the top of a rising plume or at the bottom of descending air). \remove[HW]{We will also test a}{A} common form of lateral entrainment \cite<eg.>{DBF+13} \add[HW]{is also tested in section {\protect\ref{sec:results}}}:
\begin{equation}
\varepsilon=\frac{0.2}{r_{c}}
\label{eq:fracEnt}
\end{equation}
where $r_{c}$ is the cloud or plume radius. 
\remove[HW]{Fractional entrainment is related to mass transfers by:}

@@

Next, the momentum transfers due to mass transfer and due to drag
are calculated implicitly, similar to the implicit transfers of buoyancy,
calculating $u_{i}^{n+1}$ from $u_{i}^{\prime\prime}$. These do
not influence the total $P$ or total divergence because momentum
is transferred conservatively. However they do influence $p_{i}$
which is why they are calculated before $p_{i}$. The final stage
of the time step iteration is to solve Helmholtz equations for each
$p_{i}$:
\begin{equation}
\sigma_{i}^{n+1}p_{i}^{n+1}=-\gamma\nabla\cdot\left(\sigma_{if}^{n+1}u_{i}^{n+1}-\frac{\Delta t}{2}\sigma_{if}^{n+1}\nabla_{f}\left(p_{i}^{n+1}-p_{i}^{\ell}\right)\right).\label{eq:pi_Helm}
\end{equation}
Both $p_{i}^{n+1}$ and $p_{i}^{\ell}$ appear in eqn (\ref{eq:pi_Helm})
because $u_{i}^{n+1}$ already contains contributions from $p_{i}^{\ell}$
so these are removed before updating $p_{i}^{n+1}$. No back substitution
for $u_{i}$ is done after calculating $p_{i}^{n+1}$ because this
would destroy the divergence free constraint. Therefore $p_{i}^{n+1}$
only satisfies eqn (\ref{eq:Pi_div}) exactly if the outer iterations
converge \change[HW]{and}{implying} $p_{i}^{\ell}=p_{i}^{n+1}$.

@@

\change[HW]{If there is}
{The solutions in the top row of figure {\protect\ref{fig:multiFluidBubble_500_1}} have}
no drag between fluids, no transfers, no diffusion and
the pressures of both fluids are assumed equal \add[HW]{meaning that} the multi-fluid
equations \change[HW]{are}{should be} unstable \cite{TEB19}. The simulation with ($S_{ij}=0$)
in the top row of figure \ref{fig:multiFluidBubble_500_1} actually
has small transfers to prevent either $\sigma_{i}$ from becoming
negative. This is sufficient to stabilise the solution but does not
prevent oscillations from developing in all fields. The buoyant fluid
has risen little from the initial conditions. $\sigma_1$ oscillates
between zero and one meaning that only one fluid is present at some
locations so there is no possibility for the fluids to move past each
other.
\add[HW]{The two fluids have mixed due to the two way stabilising transfers which has reduced the buoyancy in the rising fluid and increased the buoyancy in the sinking fluid.}

>>>>

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
%\usepackage{soul}
\soulregister\ref7
\soulregister\cite7
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%
\addeditor{HW}
\addeditor{WM}

@@

\citeA{TWV+18} proposed multi-fluid modelling of convection in order
to treat non-equilibrium and net mass transport \remove[HW]{correctly and} consistently.
\add[HW]{The multi-fluids approach has an advantage over ADHOC and CLUBB in solving prognostic equations for first-order moments directly (mean properties of eg. updraft and environment) rather than solving prognostic equations for higher-order moments of the whole fluid in order to diagnose means of updrafts and the environment. Solving directly for sub-fluid means could be less sensitive to closure assumptions.}
\remove{with the same numerics and physics to prognose properties of separate parts of the fluid, for example updrafts, the stable environment and downdrafts dynamics}. \citeA{Yano14} derived similar
equations \remove[HW]{for multiple segments} and showed how this is a generalisation
of the mass-flux formulation. The \citeA{TKP+18} ``extended eddy
diffusivity mass flux'' or extended EDMF scheme \remove[HW]{also} uses similar
equations but \citeA{TKP+18}
\change[HW]{didn't propose a suitable numerical method for solving the equations.}
{only solve the equations in a single column without coupling to the continuity equation of a dynamical core.}
\change[HW]{Multi-fluids promises big improvements
in the representation of sub-grid-scale convection and unification
with the representation of turbulence, particularly at the grey zone.
However, there is a long way to go, particularly as multi-fluids}
{The disadvantage of the multi-fluids approach is that it} requires
a dedicated model rather than being a stand alone parameterisation
for a dynamical core. \citeA{TEB19} and \citeA{WM19} presented solutions
of \change[HW]{the multi}{two}-fluid equations
\change[HW]{assuming two fluids that share}{using} the same pressure
\add[HW]{for both fluids}.
\remove[HW]{(or that pressure differences are predominantly due to drag
between the fluids)}. \citeA{TEB19} showed that these equations \remove[HW]{(without
drag)} are in fact unstable and can be stabilised by diffusion of vertical
velocity whereas \citeA{WM19} stabilised the equations with diffusion
and drag between fluids. Neither of these stabilisation techniques
are acceptable;
\add[HW]{sufficient diffusion to stabilise the multi-fluid equations may smooth out real features of the flow and the diffusion between fluids and drag used by \protect\citeA{WM19} meant that both fluids tended to move together rather than through each other.}
In this paper we show that there are significant
and sustained pressure differences between the fluids (section \ref{sec:results}) \add[HW]{and that the use of separate pressures stabilises the equations}.

@@

Many of the challenges of representing convection with the multi-fluid
Navier-Stokes equations \cite<e.g. those described in>{WM19} carry
over to the Boussinesq equations \add[HW]{because} convection is buoyancy dominated
and close to incompressible. The resolved test cases presented in
section \ref{sec:results} give very similar results for Navier-Stokes
and Buossinesq equations. Therefore\change[WM]{ for simplicity we focus on the}{, the}
\remove[HW]{multi-fluid} Boussinesq equations without background stratification \add[WM]{are chosen for simplicity}:

\note[HW]{Additional material until the start of 2.1}
\begin{eqnarray}
\frac{\partial \mathbf{u}}{\partial t} + \nabla\cdot(\mathbf{u}\mathbf{u})
+\nabla P & = & b \mathbf{k}+\nu\nabla^{2}\mathbf{u}
\label{eq:singleMom}
\\
\frac{\partial b}{\partial t} + \nabla\cdot(\mathbf{u}b)
 & = & \alpha\nabla^{2}b
\label{eq:singleb}
\\
\nabla\cdot\mathbf{u} & = & 0
\label{eq:singleDivFree}
\end{eqnarray}
where variables are defined in table \ref{tab:defns}.

@@

The single-fluid Boussinesq equations will be conditionally averaged following \citeA{TWV+18} to give the multi-fluid Boussinesq equations. Lagrangian labels for each fluid $i$ are defined as $I_i=1$ in fluid $i$ and 0 elsewhere. 
For the derivation of the conditionally averaged equations \change[WM]{we will not consider relabelling of fluid}{the relabelling of fluids will not be considered,} \add[HW]{and so} $I_i$ satisfies:
\begin{equation}
\frac{DI_i}{Dt}=0.
\label{eq:LagrangianLabel}
\end{equation}
\change[WM]{Applying conditional volume average to $I_i$ we define the volume fraction of fluid $i$ and the velocity and buoyancy of fluid $i$}{By applying a conditional volume average (denoted by the overbar), the fluid $i$ volume fractions, velocities and buoyancies are defined by}:
\begin{eqnarray}
\sigma_i &=& \overline{I_i} \\
\sigma_i \mathbf{u}_i &=& \overline{I_i \mathbf{u}} \\
\sigma_i b_i &=& \overline{I_i b}
\label{eq:defineFluidFields}
\end{eqnarray}
which implies
\begin{equation}
\sum_{i}\sigma_{i}  =  1.
\label{eq:sumOne}
\end{equation}
\change[WM]{As we are considering the Boussinesq equations we assume}{As the system is governed by the Boussinesq equations, it is assumed} that variations in density only affect the \change[WM]{buoyancy terms}{definition of the buoyancy terms (table \protect\ref{tab:defns})} so $\mathbf{u}_i$ and $b_i$ are defined without density weighting.

@@

Next \change[WM]{we will consider conditionally averaging the terms of }{the conditional averaging of terms in} the momentum (\ref{eq:singleMom}) and buoyancy (\ref{eq:singleb}) equations\add[WM]{ is considered}. Conditionally averaging the advection terms leads to sub-filter-scale fluxes, $F_{SF}$, since the velocity and buoyancy are not uniform within each fluid at the filter scale:
\begin{eqnarray}
\overline{I_i \nabla \cdot (\mathbf{u} \mathbf{u})} &=& \nabla \cdot (\sigma_i \mathbf{u}_i  \mathbf{u}_i) + F_\text{SF}^{\mathbf{u}_i} \\
\overline{I_i \nabla (\mathbf{u} b)} &=&  \nabla \cdot (\sigma_i \mathbf{u}_i b_i) + F_\text{SF}^{b_i}
\label{eq:filterAdvection}
\end{eqnarray}
The sub-filter-scale fluxes will be absorbed into the diffusion terms for momentum and buoyancy.
%
\note[HW]{Dan, can we do better than this?}
\change[WM]{Lots of}{Many} terms appear when attempting to conditionally average the diffusion terms which we do not know how to parameterise. \change[WM]{Therefore we will use simply}{The conditionally averaged diffusion terms are therefore assumed to behave like}:
\begin{eqnarray}
\overline{I_i \nu \nabla^2 \mathbf{u}} &=& \sigma_i \nu \nabla^2 \mathbf{u}_i \\
\overline{I_i \alpha \nabla^2 b} &=& \sigma_i \alpha \nabla^2b_i .
\end{eqnarray} 
%
\change[WM]{Unlike Thuburn et al. (2018) we define separate pressures for each fluid, $P_i$. The pressures are defined to satisfy:}{The pressures for each fluid, $P_i$, are defined to satisfy:}
\begin{equation}
\overline{I_i \nabla P} = \sigma_i \nabla P_i
\end{equation}
\add[WM]{which differs from \protect\citeA{TWV+18} who use a shared fluid pressure.}
%
Combining \change[WM]{these}{the above terms} gives the conditionally averaged momentum and buoyancy equations:
\begin{eqnarray}
\frac{\partial\sigma_i\mathbf{u}_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} \mathbf{u}_{i})
+ \sigma_i \nabla P_{i}
& = &
\sigma_i b_{i}\mathbf{k}
+
\nu\sigma_i\nabla^{2}\mathbf{u}_{i}
+
\sum_{j\ne i}\left\{ \sigma_{j}S_{ji}\mathbf{u}_{ji}^{T}
 - \sigma_{i}\mathbf{D}_{ij}\right\}
\label{eq:momFlux}\\
\frac{\partial\sigma_i b_{i}}{\partial t} + 
\nabla\cdot(\sigma_i \mathbf{u}_{i} b_{i})
& = &
\alpha\sigma_i\nabla^{2}b_{i}
+
\sum_{j\ne i} \sigma_{j}S_{ji} b_{ji}^{T}.
\label{eq:bFlux}
\end{eqnarray}
These equations include mass transfers between fluids with $\mathbf{u}_{ij}^T$ being the velocity of the fluid transferred from $i$ to $j$ and $b_{ij}^T$ being the buoyancy of the fluid transferred from $i$ to $j$. A drag term, $\mathbf{D}_{ij}$ is also included to parameterise some of the missing terms because it is commonly used when conditional averaging is applied in fields such as multi-phase modelling \cite<eg>{GBB+07}. Sensitivity to $\mathbf{D}_{ij}$ will be investigated in section \ref{sec:results}.

@@

\change[WM]{We will assume that}{It is assumed that the} pressure in each fluid is given by a \remove[HW]{parameterised}
perturbation, $p_{i}$, from the total pressure, $P$:
\begin{equation}
P_{i}=P+p_{i}.
\end{equation}
\add[HW]{
For stability and respect $\sum_i \nabla\cdot\sigma_i\mathbf{u}_i=0$, we propose four constraints on the pressure perturbations, $p_i$:}
\begin{enumerate}
\item $\sum_{i}\sigma_{i}p_{i}=0$.
\item $\mathbf{u}_{i}=\mathbf{u} \implies p_i=0$ so that pressure perturbations do not force individual fluid anomalies away from the mean.
\item $p_i \not\to 0$ when $\sigma_i\to 0$ so that pressure perturbations control divergence in vanishing fluids.
\item $p_i\not\to\infty$ as $\sigma_j\to 0$ for any combination of $i,j$.
\end{enumerate}

\add[HW]{In order to derive a suitable form of parameterisation for $p_i$ we will first consider the} \citeA{BN86} \add{model of multi-phase flow.}

@@

\citeA{BN86} proposed a parameterisation for pressure differences
in a multi-phase explosion model that has been widely used in mathematics
and engineering for simulating, \add[HW]{for example}, fluidised beds of granular material \cite<e.g.>{EHM92} and two compressible fluids \cite<e.g.>{SA99}. \add[HW]{This model is not used here but is examined in order to derive a parameterisation of pressure differences between fluids suitable for atmospheric convection.}
\citeA{SA99} wrote transport equations for both the volume fraction,
$\sigma_{i}$, and for the mass fraction,
$\sigma_{i}\rho_{i}$  without transfers between fluids which in our notation are:
\begin{eqnarray}
\frac{\partial\sigma_{i}}{\partial t}+\mathbf{V}\cdot\nabla\sigma_{i} & = & \mu\left(P_{i}-P_{j}\right)
\label{eq:BNcompaction}
\\
\frac{\partial\sigma_{i}\rho_{i}}{\partial t}+\nabla\cdot\left(\sigma_{i}\rho_{i}\mathbf{u}_{i}\right) & = & 0
\label{eq:rhoTransport}
\end{eqnarray}
where $\mathbf{V}$ is the velocity of the interface between fluids,
$\mu$ is the compaction viscosity which controls how quickly the
pressure of each fluid relaxes to the mean pressure and index $j$
refers to the other fluid in a two fluid system.
\remove[HW]{In our Boussinesq system, $\rho_{i}$ does not appear in the continuity equation.}
Different authors make different assumptions about the interface velocity.
\change[HW]{but if we assume that $\mathbf{V}=\sum_{i}\sigma_{i}\mathbf{u}_{i}$ and $\rho_{i}$ is constant then eqns (\ref{eq:divFree}), (\ref{eq:sumOne}), (\ref{eq:BNcompaction}) and (\ref{eq:rhoTransport}) imply that $\mu p_{i}=-\sigma_{i}^{2}\nabla\cdot\mathbf{u}_{i}$. If we assume that $\mathbf{V}=\mathbf{u}_{i}$ for each fluid then we get $\mu p_{i}=-\sigma_{i}\nabla\cdot\mathbf{u}_{i}$.}
{The combination of equations (\ref{eq:BNcompaction}) and (\ref{eq:rhoTransport}) shows that the pressure difference between fluids is related to divergence.}

@@

\citeA{TEB19} stabilised the multi-fluid equations by adding a diffusion
term $\nu\partial w^{2}/\partial z^{2}$ to the RHS of the $w$ equation.
If a fluid perturbation pressure is set to \change[WM]{$-\nu\partial w/\partial z$}{$p_i=-\nu\partial w/\partial z$}
\add[WM]{then} this is equivalent to the diffusion term in one dimension. However
adding a diffusion term will harm three dimensional simulations of
resolved convection in the free troposphere whereas adding a fluid
perturbation pressure will not. Inspired by both \citeA{BN86} and
\citeA{TEB19}
\change[WM]{we assume that the pressure perturbation in each fluid is given
by}{and obeying the constraints given above, the pressure in each fluid is chosen to be}:
\begin{equation}
p_{i}=\gamma\sum_j \sigma_j\nabla\cdot\mathbf{u}_j -\gamma\nabla\cdot\mathbf{u}_{i}
\label{eq:Pi_div}
\end{equation}
where $\gamma$ is a \change[HW]{free coefficient}{compressibility with units $\text{m}^s \text{s}^{-1}$. It should be noted that this parameterisation destroys energy conservation.}

@@

\add[HW]{The transfers between fluids are related to entrainment, $\varepsilon$, and detrainment, $\delta$. Fluid 0 is the environment and fluid 1 is updrafts implying:}
\begin{equation}
S_{01} = w_1\varepsilon, \;\;\;\;\;\;\;
S_{10} = w_0\delta.
\end{equation}
\add[HW]{Results using two formulations of entrainment are presented in section {\protect\ref{sec:results}}. The first is similar to}
\remove[HW]{We consider dry air without turbulence so use}
the dynamic entrainment described by
\citeA{HC51}, \citeA{AK67} and \citeA{DBF+13}:
\begin{equation}
S_{ij}=\begin{cases}
-\nabla\cdot\mathbf{u}_{i} & \text{if }\nabla\cdot\mathbf{u}_{i}<0\\
0 & \text{otherwise.}
\end{cases}
\label{eq:Sdiv}
\end{equation}
\change[HW]{which is necessary to stop air in}{This formulation prevents} one fluid from accumulating when \change[WM]{vertical motion ceases}{a fluid decelerates} (at the top of a rising plume or at the bottom of descending air). \remove[HW]{We will also test a}{A} common form of lateral entrainment \cite<eg.>{DBF+13} \add[HW]{is also tested in section {\protect\ref{sec:results}}}:
\begin{equation}
\varepsilon=\frac{0.2}{r_{c}}
\label{eq:fracEnt}
\end{equation}
where $r_{c}$ is the cloud or plume radius. 
\remove[HW]{Fractional entrainment is related to mass transfers by:}

@@

Next, the momentum transfers due to mass transfer and due to drag
are calculated implicitly, similar to the implicit transfers of buoyancy,
calculating $u_{i}^{n+1}$ from $u_{i}^{\prime\prime}$. These do
not influence the total $P$ or total divergence because momentum
is transferred conservatively. However they do influence $p_{i}$
which is why they are calculated before $p_{i}$. The final stage
of the time step iteration is to solve Helmholtz equations implicitly for each
$p_{i}^{n+1}$:
\begin{equation}
p_{i}^{n+1} = \gamma \sum_j \left(\sigma^{n+1}_j \nabla\cdot u_j^{n+1} \right)
-\gamma\nabla\cdot
\left(u_{i}^{n+1}-\frac{\Delta t}{2}\nabla_{f}\left(p_{i}^{n+1}-p_{i}^{\ell}\right)\right)
\label{eq:pi_Helm}
\end{equation}
Both $p_{i}^{n+1}$ and $p_{i}^{\ell}$ appear in eqn (\ref{eq:pi_Helm})
because $u_{i}^{n+1}$ already contains contributions from $p_{i}^{\ell}$
so these are removed before updating $p_{i}^{n+1}$. No back substitution
for $u_{i}$ is done after calculating $p_{i}^{n+1}$ because this
would destroy the divergence free constraint. Therefore $p_{i}^{n+1}$
only satisfies eqn (\ref{eq:Pi_div}) exactly if the outer iterations
converge \change[HW]{and}{implying} $p_{i}^{\ell}=p_{i}^{n+1}$.

@@

\change[HW]{If there is}
{The solutions in the top row of figure {\protect\ref{fig:multiFluidBubble_500_1}} have}
no drag between fluids, no transfers, no diffusion and
the pressures of both fluids are assumed equal \add[HW]{meaning that} the multi-fluid
equations \change[HW]{are}{should be} unstable \cite{TEB19}. The simulation with ($S_{ij}=0$)
in the top row of figure \ref{fig:multiFluidBubble_500_1} actually
has small transfers to prevent either $\sigma_{i}$ from becoming
negative. This is sufficient to stabilise the solution but does not
prevent oscillations from developing in all fields. The buoyant fluid
has risen little from the initial conditions. $\sigma_1$ oscillates
between zero and one meaning that only one fluid is present at some
locations so there is no possibility for the fluids to move past each
other.
\add[HW]{The two fluids have mixed due to the two way stabilising transfers which have reduced the buoyancy in the rising fluid and increased the buoyancy in the sinking fluid.}
